\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-09-17 submission/manuscript.Rmd
Update manuscript.Rmd

In the second part of this study, 109 photos of neutral-expression faces of real individuals from the American Multiracial Face Database (AMFD, [@chen2021broadening]) were presented to the participants for 500ms each, in a randomized order. Following each stimulus, ratings of **Attractiveness** ("I find this person attractive"), **Beauty** ("This face is good-looking"), **Trustworthiness** ("I find this person trustworthy") and **Familiarity** ("This person reminds me of someone I know") were collected using visual analog scales. (**TODO: add details and justifications**)

@@

This study aimed at investigating the effect of facial ratings (attractiveness, beauty, trustworthiness and familiarity) on simulation monitoring, i.e., on the belief that the stimulus is artificially generated. The most striking result is that despite all the images being real faces retrieved from the same validated database, all participants, when given the information, believed (to high degrees of confidence) that a significant proportion of them were fake. This result is a testimony to both the current expectations regarding CGI technology in the population, as well as to the volatility of our sense of reality. It underlines the strong impact of prior expectations and information on reality beliefs. In fact, stimuli-related and participant-related characteristics accounted for less than 20% of the beliefs variance, suggesting that a large part of it is associated with other subjective processes.

@@

Although attractiveness did not seem to be the primary drive underlying simulation monitoring of face images, we do nonetheless report significant associations, with a different pattern depending on the participant's gender. The quadratic relationship found for female participants is aligned with our hypothesis that faces which are highly salient (i.e., perceived to be very attractive or very unattractive) are judged to be more real. Alternatively, we could interpret the mostly positive linear relationship for males under an evolutionary lens.

In particular, males are widely postulated to place more importance toward facial attractiveness as a sign of reproductive potential, relative to females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender]. As such, males could have perceived highly attractive and beautiful faces to be more real since such attractiveness act as salient cues of biological quality [@fink2006facial]. Therefore, our study suggests a more complex relationship between stimulus attractiveness, demographic variables and perceptions of reality that has not been examined in the literature.

Moreover, beyond its impacts on individuals' beliefs about realness, attractiveness was found to significantly affect the confidence in judgements for females.

>>>>

In the second part of this study, 109 photos of neutral-expression faces of real individuals from the validated American Multiracial Face Database (AMFD, [@chen2021broadening]) were presented to the participants for 500ms each, in a randomized order. Following each stimulus, ratings of **Attractiveness** ("I find this person attractive"), **Beauty** ("This face is good-looking"), **Trustworthiness** ("I find this person trustworthy") and **Familiarity** ("This person reminds me of someone I know") were collected using visual analog scales. (**TODO: add details and justifications**)

@@

This study aimed at investigating the effect of facial ratings (attractiveness, beauty, trustworthiness and familiarity) on simulation monitoring, i.e., on the belief that the stimulus is artificially generated. The most striking result, in our opinion, is that despite all the stimuli being real faces from the same database, all participants, when given the information, believed (to high degrees of confidence) that a significant proportion of them were fake. This finding is a testimony to both the current expectations regarding CGI technology in the population, as well as to the volatility of our sense of reality. It underlines the strong impact of prior expectations and information on reality beliefs. In fact, stimuli-related and participant-related characteristics accounted for less than 20% of the beliefs variance, suggesting that a large part of it is associated with other subjective processes.

@@

Although attractiveness did not seem to be the primary drive underlying simulation monitoring of face images, we do nonetheless report significant associations, with a different pattern depending on the participant's gender. The quadratic relationship found for female participants is aligned with our hypothesis that salient faces (i.e., rated as very attractive or very unattractive) are judged to be more real. Alternatively, we could interpret the mostly positive linear relationship for males under an evolutionary lens. In particular, males purportedly place more emphasis on facial attractiveness as a sign of reproductive potential, as compared with females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender; @fink2006facial]. As such, it is possible that the evolutionary weight associated with attractiveness skewed the perceived saliency towards attractive faces, rendering attractive faces are significantly more salient than unattractive ones, in turn distorting the relationship with simulation monitoring. Future studies should test this saliency-based hypothesis by measuring constructs closer to salience and its effects, for instance using neuroimaging (**ref on saliency in EEG or fMRI**) or physiological markers [in particular, heart rate deceleration, **include refs**].

<!-- Beauty -->
discuss why the link with beauty is slightly different.
